Sleep Quality Estimation using Accelerometer Data from Thigh-Mounted Devices During in Free Living Conditions

### 2023-03-16 presentation/index.qmd
working on correct tie breaking
setup for three workflows: crude, binary relevance, and multiclass
create roc data for all approaches
work on roc curves

- data preparation, big time-consumer is handling raw acc data
- only thigh data used. HSBC and other is only thigh data...
- all zm recording is considered as in-bed (sensor problem?)
- no sleep stages, only sleep/awake
- sensor problems during sleep, up to 20 consecutive epochs (200 sec) are treated as sleep
- Combine binary classifiers to produce multilabel outcomes (i.e., binary Relevance)
- Evaluated combined classifiers using weighted micro-averaging.

@@

@@

plot_ex
```

@@

plot_lps
```

::: aside 
This methedology is called binary relevance.
:::

@@

plot_tst
```

@@

## ROC Curves {transition="convex"}

```{r}
source(here::here("presentation/code/plot_roc_curves.R"))

roc_plots
```

## ROC Curves {transition="convex"}

```{r}
roc_combined_plots
```

## Lots of Metrics {transition="convex" .H2}

::: {.column}
```{r}
source(here::here("presentation/code/tables_performance.R"))

tab1
```
:::

::: {.column}
```{r}
tab2
```
:::

## Bland-Altman Plots {transition="convex"}

::: {.panel-tabset}

### Decision Tree

```{r}
#| echo: false
source(here::here("presentation/code/plot_ba.R"))

dc_plots[[1]]

dc_plots[[2]]

dc_plots[[3]]

dc_plots[[4]]

dc_plots[[5]]
```

### Logistic Regression

```{r}
lr_plots[[1]]

lr_plots[[2]]

lr_plots[[3]]

lr_plots[[4]]

lr_plots[[5]]
```

### Neural Net

```{r}
nn_plots[[1]]

nn_plots[[2]]

nn_plots[[3]]

nn_plots[[4]]

nn_plots[[5]]
```

### XGBoost

```{r}
xg_plots[[1]]

xg_plots[[2]]

xg_plots[[3]]

xg_plots[[4]]

xg_plots[[5]]
```

## Bland-Altman Analysis {transition="convex" .H2}

```{r}
source(here::here("presentation/code/table_ba.R"))

tab_ba
```

## In-bed classification flow {transition="convex" .H2}

```{r}
in_bed_tree <-
  read_rds(here::here("data/models/fitted_models/in_bed_simple_CART_fit.rds"))

in_bed_tree |>
  extract_fit_engine() |>
  visTree(
    height = "550px",
    nodesPopSize = TRUE,
    minNodeSize = 10,
    maxNodeSize = 30,
    fallenLeaves = TRUE,
    legend = FALSE,
    digits = 3,
    direction = "UD",
    colorVar = c(
      "#CC5500", "#8B0000", "#FFD700",
      "#228B22", "#654321", "#8B0000", "#FFDB58"
    ),
    export = FALSE
  ) |>
  visNodes(font = list(
    color = "#EEE8D5",
    size = 20,
    strokeWidth = 1
  )) |>
  visEdges(font = list(
    color = "#EEE8D5",
    size = 20,
    strokeWidth = 1
  )) |>
  visInteraction(
    dragNodes = FALSE,
    dragView = FALSE,
    zoomView = FALSE
  )
```

## Sleep classification flow {transition="convex" .H2}

```{r}
sleep_tree <-
  read_rds(here::here("data/models/fitted_models/sleep_simple_CART_fit.rds"))

sleep_tree |>
  extract_fit_engine() |>
  visTree(
    height = "550px",
    nodesPopSize = TRUE,
    minNodeSize = 10,
    maxNodeSize = 30,
    fallenLeaves = TRUE,
    legend = FALSE,
    digits = 3,
    direction = "UD",
    colorVar = c(
      "#CC5500", "#8B0000", "#FFD700",
      "#228B22", "#654321", "#8B0000", "#FFDB58"
    ),
    export = FALSE
  ) |>
  visNodes(font = list(
    color = "#EEE8D5",
    strokeWidth = 1
  )) |>
  visEdges(font = list(
    color = "#EEE8D5",
    strokeWidth = 1
  )) |>
  visInteraction(
    dragNodes = FALSE,
    dragView = FALSE,
    zoomView = FALSE
  )
```

## Discussion {transition="convex"}

- heteroscedasticity
- Cheung 2018 table 4: actigraphy provides a sufficiently narrow range of possible mean differences (CI 95%) clinical significant thresholds
- could be interesting to build models on thigh and hip ocmbined.
- multiclass vs multilabel classification
- in-bed awake/sleep is highly imbalanced -> maybe train a new classifier accounting for imbalanced data (SMOTE)
- model combined preds instead?

## References {transition="convex"}

>>>>

- data preparation, big time-consumer is handling raw acc data
- only thigh data used. HSBC and other is only thigh data...
- all zm recording is considered as in-bed (sensor problem?)
- no sleep stages, only sleep/awake
- sensor problems during sleep, up to 20 consecutive epochs (200 sec) are treated as sleep
- Combine binary classifiers to produce multilabel outcomes (i.e., binary Relevance)
- Evaluated combined classifiers using weighted micro-averaging.
- adjust threshold for in bed awake or give edge in tie breakers??
- zm sleep stats inclusion criteria
- three approaches: tw- binary (risk of ghost classes), binary relevance (choosing the right tie breaker), and multiclass

@@

## Present three modeling approaches

::: {.fragment}
- Multiclass classifier with three labels: in-bed-asleep, in-bed-awake, and out-bed-awake
  - pros: simplicity, speed, joint modeling, consistency across classes
  - cons: struggle with class imbalance, performance
:::

::: {.fragment}
- Two binary classifiers: in-bed/out-bed and asleep/awake
  - pros: 
  - cons: ghost classes
:::

::: {.fragment}
- Three binary classifiers: in-bed-asleep 1/0, in-bed-awake 1/0, and out-bed-awake 1/0
  - pros: more flexible, can better handle class imbalance
  - cons: complex training, low speed, feature redundancy
:::

@@

multiclass_ex
```

@@

```{r}
#| fig-asp: .4
multiclass_lps
```

@@

```{r}
#| fig-asp: .4
multiclass_tst
```

```{r}
#| fig-asp: .4
plot_ex
```

```{r}
#| fig-asp: .4
plot_lps
```

```{r}
#| fig-asp: .4
plot_tst
```

@@

<!-- ## ROC Curves {transition="convex"} -->
<!-- ### Two binary models (in-bed/asleep) -->

<!-- ```{r} -->
<!-- source(here::here("presentation/code/plot_roc_curves.R")) -->

<!-- roc_plots -->
<!-- ``` -->

<!-- ## ROC Curves {transition="convex"} -->
<!-- ### Binary relevance -->

<!-- ```{r} -->
<!-- source(here::here) -->
<!-- ``` -->

<!-- ## Lots of Metrics {transition="convex" .H2} -->

<!-- :::: {.columns} -->

<!-- ::: {.column} -->
<!-- ```{r} -->
<!-- source(here::here("presentation/code/tables_performance.R")) -->

<!-- tab1 -->
<!-- ``` -->
<!-- ::: -->

<!-- ::: {.column} -->
<!-- ```{r} -->
<!-- tab2 -->
<!-- ``` -->
<!-- ::: -->

<!-- :::: -->

<!-- ## Bland-Altman Plots {transition="convex"} -->

<!-- ::: {.panel-tabset} -->

<!-- ### Decision Tree -->

<!-- ```{r} -->
<!-- #| echo: false -->
<!-- source(here::here("presentation/code/plot_ba.R")) -->

<!-- dc_plots[[1]] -->

<!-- dc_plots[[2]] -->

<!-- dc_plots[[3]] -->

<!-- dc_plots[[4]] -->

<!-- dc_plots[[5]] -->
<!-- ``` -->

<!-- ### Logistic Regression -->

<!-- ```{r} -->
<!-- lr_plots[[1]] -->

<!-- lr_plots[[2]] -->

<!-- lr_plots[[3]] -->

<!-- lr_plots[[4]] -->

<!-- lr_plots[[5]] -->
<!-- ``` -->

<!-- ### Neural Net -->

<!-- ```{r} -->
<!-- nn_plots[[1]] -->

<!-- nn_plots[[2]] -->

<!-- nn_plots[[3]] -->

<!-- nn_plots[[4]] -->

<!-- nn_plots[[5]] -->
<!-- ``` -->

<!-- ### XGBoost -->

<!-- ```{r} -->
<!-- xg_plots[[1]] -->

<!-- xg_plots[[2]] -->

<!-- xg_plots[[3]] -->

<!-- xg_plots[[4]] -->

<!-- xg_plots[[5]] -->
<!-- ``` -->

<!-- ::: -->

<!-- ## Bland-Altman Analysis {transition="convex" .H2} -->

<!-- ```{r} -->
<!-- source(here::here("presentation/code/table_ba.R")) -->

<!-- tab_ba -->
<!-- ``` -->

<!-- ## In-bed classification flow {transition="convex" .H2} -->

<!-- ```{r} -->
<!-- in_bed_tree <- -->
<!--   read_rds(here::here("data/models/fitted_models/in_bed_simple_CART_fit.rds")) -->

<!-- in_bed_tree |> -->
<!--   extract_fit_engine() |> -->
<!--   visTree( -->
<!--     height = "550px", -->
<!--     nodesPopSize = TRUE, -->
<!--     minNodeSize = 10, -->
<!--     maxNodeSize = 30, -->
<!--     fallenLeaves = TRUE, -->
<!--     legend = FALSE, -->
<!--     digits = 3, -->
<!--     direction = "UD", -->
<!--     colorVar = c( -->
<!--       "#CC5500", "#8B0000", "#FFD700", -->
<!--       "#228B22", "#654321", "#8B0000", "#FFDB58" -->
<!--     ), -->
<!--     export = FALSE -->
<!--   ) |> -->
<!--   visNodes(font = list( -->
<!--     color = "#EEE8D5", -->
<!--     size = 20, -->
<!--     strokeWidth = 1 -->
<!--   )) |> -->
<!--   visEdges(font = list( -->
<!--     color = "#EEE8D5", -->
<!--     size = 20, -->
<!--     strokeWidth = 1 -->
<!--   )) |> -->
<!--   visInteraction( -->
<!--     dragNodes = FALSE, -->
<!--     dragView = FALSE, -->
<!--     zoomView = FALSE -->
<!--   ) -->
<!-- ``` -->

<!-- --- -->

<!-- ## Sleep classification flow {transition="convex" .H2} -->

<!-- ```{r} -->
<!-- sleep_tree <- -->
<!--   read_rds(here::here("data/models/fitted_models/sleep_simple_CART_fit.rds")) -->

<!-- sleep_tree |> -->
<!--   extract_fit_engine() |> -->
<!--   visTree( -->
<!--     height = "550px", -->
<!--     nodesPopSize = TRUE, -->
<!--     minNodeSize = 10, -->
<!--     maxNodeSize = 30, -->
<!--     fallenLeaves = TRUE, -->
<!--     legend = FALSE, -->
<!--     digits = 3, -->
<!--     direction = "UD", -->
<!--     colorVar = c( -->
<!--       "#CC5500", "#8B0000", "#FFD700", -->
<!--       "#228B22", "#654321", "#8B0000", "#FFDB58" -->
<!--     ), -->
<!--     export = FALSE -->
<!--   ) |> -->
<!--   visNodes(font = list( -->
<!--     color = "#EEE8D5", -->
<!--     strokeWidth = 1 -->
<!--   )) |> -->
<!--   visEdges(font = list( -->
<!--     color = "#EEE8D5", -->
<!--     strokeWidth = 1 -->
<!--   )) |> -->
<!--   visInteraction( -->
<!--     dragNodes = FALSE, -->
<!--     dragView = FALSE, -->
<!--     zoomView = FALSE -->
<!--   ) -->
<!-- ``` -->

<!-- ## Discussion {transition="convex"} -->

<!-- - heteroscedasticity -->
<!-- - Cheung 2018 table 4: actigraphy provides a sufficiently narrow range of possible mean differences (CI 95%) clinical significant thresholds -->
<!-- - could be interesting to build models on thigh and hip ocmbined. -->
<!-- - multiclass vs multilabel classification -->
<!-- - in-bed awake/sleep is highly imbalanced -> maybe train a new classifier accounting for imbalanced data (SMOTE) -->
<!-- - model combined preds instead? -->

<!-- ## References {transition="convex"} -->
